Benefits and risks associated with children’s and adolescents’ interactions with electronic screens: An umbrella review

### 2023-03-12 reports/manuscript.Rmd
Address some coauthor comments

In 2016, the American Academy of Pediatrics used a narrative review to examine the benefits and risks of children and adolescents' electronic media [@chassiakosChildrenAdolescentsDigital2016] as a basis for updating their guidelines about screen use [@AAPMediaUseSchoolAged2016].
Since then, a large number of systematic reviews and meta-analyses have provided evidence about the potential benefits and risks of screen use.
<span id = "r2_9">
While there have been other overviews of reviews on screen time, these have tended to focus on a single domain (e.g., health [@stiglicEffectsScreentimeHealth2019]), focus on a particular exposure (e.g., social media [@valkenburgSocialMediaUse2022;@arias-delatorreRelationshipDepressionUse2020]) or provide only a narrative summary of the literature [@orbenTeenagersScreensSocial2020].
<span id = "r1_2">No review has yet examined the evidence available across a broad range of outcome domains, such as physical health, education, physical and cognitive development, behaviour, and well-being.
By summarising and synthesising all evidence in one overview, we provide a reference point for the field and allow for easier comparison of risks and benefits for the same behaviour.</span></span>

@@

Largely owing to a small number of studies or missing individual study data, there were few age-based conclusions that could be drawn from reviews which met our criteria for statistical certainty.
If we expand to include those reviews which did not meet this threshold, there remained no clear pattern although there were some age-specific differences in associations (data available in Supplementary Materials).
For example, advertising of unhealthy food was associated with unhealthy food choice for young children, but was not statistically significant for other age groups [@sadeghiradInfluenceUnhealthyFood2016].
Conversely, TV programs and movies were more strongly associated with lower physical activity for adolescents than for younger age groups [@marshallRelationshipsMediaUse2004].

@@

<span id="r1_4">Our high-level approach also means that we could not engage with the specific mechanisms behind each association, and as such, we cannot comment on the evidence for causality.
Instead, readers who wish to more deeply understand one specific relationship are directed to the cited review for that effect, where the authors could engage more deeply with the mechanisms.</span>
<span id="r2_16">We converted all effect sizes to a common metric (Pearson's r) to allow for comparisons of magnitude, but acknowledge that this assumes a linear relationship.
Some previous research suggests that associations are typically linear [@sandersTypeScreenTime2019].
However, others have identified instances where non-linear relationships exist, especially for very high levels of screen time [@twengeMoreTimeTechnology2019; @przybylskiLargeScaleTestGoldilocks2017; @kellySocialMediaUse2018].
Additionally, our conversion may not always adequately account for differences in study design or measures of exposures and outcomes.
Care is needed, therefore, when interpreting the effect sizes.</span>
<span id = "r3_2_2">In addition, reviews provide only historical evidence which may not keep up with the changing ways children can engage with screens.
While our synthesis of the existing evidence provides information about how screens might have influenced children in the past, it is difficult to know if these findings will translate to new forms of technology in the future.</span>

>>>>

In 2016, the American Academy of Pediatrics used a narrative review to examine the benefits and risks of children and adolescents' electronic media [@chassiakosChildrenAdolescentsDigital2016] as a basis for updating their guidelines about screen use [@AAPMediaUseSchoolAged2016].
Since then, a large number of systematic reviews and meta-analyses have provided evidence about the potential benefits and risks of screen use.
<span id = "r2_9">
While there have been other overviews of reviews on screen time, these have tended to focus on a single domain (e.g., health [@stiglicEffectsScreentimeHealth2019]), focus on a particular exposure (e.g., social media [@valkenburgSocialMediaUse2022;@arias-delatorreRelationshipDepressionUse2020]) or provide only a narrative summary of the literature [@orbenTeenagersScreensSocial2020].
<span id = "r1_2">
Focusing on a single domain or exposure makes it difficult to understand what trade-offs are involved in any guidelines around screen use.
For example, prohibiting screen use might reduce exposure to advertising but may also thwart learning opportunities from interactive educational tools. Reviews on either of these exposures or outcomes would likely miss being able to quantify these trade-offs.
Overviews are one method of evidence synthesis that helps address these trade-offs, by providing 'user-friendly' summaries of a field of research [@pollockChapterOverviewsReviews2022].
These overviews provide a reference point for the field and allow for easier comparison of risks and benefits for the same behaviour.</span></span>

@@

<span id = "r1_5">
Largely owing to a small number of studies or missing individual study data, there were few age-based conclusions that could be drawn from reviews which met our criteria for statistical certainty.
If we expand to include those reviews which did not meet this threshold, there remained no clear pattern although there were some age-specific differences in associations (data available in Supplementary Materials).
For example, advertising of unhealthy food was associated with unhealthy food choice for young children, but was not statistically significant for other age groups [@sadeghiradInfluenceUnhealthyFood2016].
Conversely, TV programs and movies were more strongly associated with lower physical activity for adolescents than for younger age groups [@marshallRelationshipsMediaUse2004].
Given the differences in development across childhood and adolescence and the different ways children of various ages use screens, further examination of age-based differences is needed.
However, in the absence of this work, our study has shown how children are affected by screens in general.
</span>

@@

<span id="r1_4">Our high-level approach also means that we could not engage with the specific mechanisms behind each association, and as such, we cannot make strong claims on the directions of causality.
These likely depend on the specific exposure and outcome.
It is tempting for people to draw inferences that the associations are due to screen time causing these outcomes, but we cannot rule out reverse causality, a third variable, or some combination of influences.
Many of the individual reviews go into more detail about the strength of the evidence for causal associations, but those judgements were difficult to synthesise across more than 200 reviews.
Readers who wish to more deeply understand one specific relationship are directed to the cited review for that effect, where the authors could engage more deeply with the mechanisms.</span>
<span id="r2_16">We converted all effect sizes to a common metric (Pearson's r) to allow for comparisons of magnitude, but acknowledge that this assumes a linear relationship.
Some previous research suggests that associations are typically linear [@sandersTypeScreenTime2019].
However, others have identified instances where non-linear relationships exist, especially for very high levels of screen time [@twengeMoreTimeTechnology2019; @przybylskiLargeScaleTestGoldilocks2017; @kellySocialMediaUse2018].
Additionally, our conversion may not always adequately account for differences in study design or measures of exposures and outcomes.
Care is needed, therefore, when interpreting the effect sizes.</span>
<span id = "r3_2_2">In addition, reviews provide only historical evidence which may not keep up with the changing ways children can engage with screens.
While our synthesis of the existing evidence provides information about how screens might have influenced children in the past, it is difficult to know if these findings will translate to new forms of technology in the future.</span>

### 2023-03-12 reports/revision_letter.Rmd
Address some coauthor comments

Please see our response to Reviewer 2, in particular \comment{r2_29}.
The reviewer raised some important concerns, especially regarding odds ratios which we have decided should not have been included.
For the other issues, we have provided stronger justification for our decisions.

@@

We agree with the reviewer that the term 'screen time' is problematic. 
Indeed, as our results show, the influence of screen time on children is a less a function of a physical screen, and more a function of the activity that the screen provides. 
But, we feel that a review such as this is needed in order to emphasise the diverse influence that screens can have. 
If the results are not collated into a single place, it is difficult to see how the field will begin to progress away from 'screen time' as a concept. 
We have included this in the *Implications for Future Research* section:

@@

Our goal with this review was to provide an overview of the field of screen time research.
A summary of an entire field will, out of necessity, be more broad that a typical systematic review.
We believe that there is significant value in providing a single overview to summarise the state of the field.
We have added some additional information to the introduction on this point:

@@

We intended to include the review which captured the most information on the exposure/outcome combination for each effect size.
The assumption was that the largest review would be the most recent, and that he newer reviews would contain the studies located in previous reviews. 
Manual inspection of the reviews which did not contribute any effects suggests that this was successful - older reviews tended to only be included if they focused on the general population while a newer review focused on a sub-population.

@@

Yes, we agree that the body of evidence does not lend itself to strong causal conclusions.
We note that we deliberately chose to avoid using strong causal language as much as possible, instead indicating that the evidence suggests associations with outcomes.
We have reread the manuscript and removed any remaining causal language that we found.
We have also added the following to the *Discussion* section:

@@

Note that we do use causal language when describing hypothesised mechanisms, and when describing guidelines.
We feel that this is sensible --- guidelines are based on causal assumptions, and while our evidence does not always support them, we are referring to them in the context that they are made (i.e., to instruct parents).

@@

One of the benefits of the other reviewers' and editor's suggestion to update the search is that some of this limitation has been addressed in more recent literature.
Newer papers were more likely to provide moderation by age group, or new reviews targeting specific age groups were now available.
Unfortunately, many of these still failed to meet our criteria for statistical certainty.
Inspecting the results in the supplementary material still reveal fewer differences by age group than may be expected.
In some cases this is because there is still only an effect for a single age group, as many reviews target behaviours or outcomes which are more common in one age group (e.g., sexting in adolescents, or developmental outcome in young children).
For the effects where there were multiple age groups, differences tended to be small (e.g., body composition).

@@

The quote from the *Lancet* editorial is that the understanding is lacking, rather than the evidence.
Our understanding is lacking partly because benefits of screen time (as seen in some educational literature) and reviewed separately from harms.
Our review aims to address this.
See additional context in the *Introduction*:

@@

We have revised the introduction with new papers, with a particular focus on other umbrella reviews.

>>>>

Please see our response to Reviewer 2, in particular \comment{r2_29}.
The reviewer raised some important concerns, especially regarding the interpretability of odds ratios which we have decided should not have been included (see \comment{r2_16}).
For the other issues, we have provided stronger justification for our decisions.

@@

We agree with the reviewer that the term 'screen time' is problematic as it fails to acknowledge the diversity of experiences screens provide.
Indeed, our motivation for conducting this review was to counter this narrative in the literature.
To do this, it was necessary to invoke the concept of 'screen time'.
As our results show, the influence of screen time on children is less a function of a physical screen, and more a function of the activity that the screen provides.
But, we feel that a review such as this is needed in order to emphasise the diverse influence that screens can have. 
If the results are not collated into a single place, it is difficult to see how the field will begin to progress away from 'screen time' as a concept. 
We have included this in the *Implications for Future Research* section:

@@

We acknowledge that the field contains a range of different effects with different magnitudes and directions.
We want to make clear what trade-offs are involved in any guidelines around screen use.
For example, prohibiting screen use might reduce exposure to advertising but may also thwart learning opportunities from interactive educational tools. Reviews on either of these exposures or outcomes would likely miss being able to quantify these trade-offs.
As a result, we think casting a wide net and using a meta-review methodology is one of the few ways of pragmatically outlining and quantifying the trade-offs.
As Cochrane says, "Overviews are an increasingly popular form of evidence synthesis, as they aim to provide 'user-friendly' summaries of the breadth of research relevant to a decision without decision makers needing to assimilate the results of multiple systematic reviews themselves (https://training.cochrane.org/handbook/current/chapter-v).

@@

We intended to include the review which captured the most information on the exposure/outcome combination for each effect size.
The assumption was that the largest review would be the most recent, and that newer reviews would contain the studies located in previous reviews. 
We inspected the reviews that did not contribute any effects.
This inspection suggested our assumptions were correct: older reviews tended to only be included if they focused on the general population while a newer review focused on a specific sub-population.

@@

Yes, we agree that the body of evidence does not lend itself to strong causal conclusions.
A strength of our study is that it points out issues with how research has been conducted in this field, which we hope will make future research more ambitious.
We note that we deliberately chose to avoid using strong causal language as much as possible, instead indicating that the evidence suggests associations with outcomes.
We have carefully reread the manuscript and removed any remaining unjustified causal language.
We have also added the following to the *Discussion* section:

@@

Note that we do use causal language when describing hypothesised mechanisms, and when describing guidelines.
We feel that this is sensible---guidelines are based on causal assumptions, and while our evidence does not always support them, we are referring to them in the context that they are made (i.e., to instruct parents).

@@

We agree that it would be useful to break down findings by age.
However, for us to accomplish this we need meta-analyses which examine the developmental span for each exposure and outcome combination.
While the addition of further reviews based on the reviewers' and editor's suggestion to update the search helped address this limiation (ewer papers were more likely to provide moderation by age group, and new reviews targeting specific age groups have been added), there were still too few age-based comparisons to make strong claims.
This is because many reviews target behaviours or outcomes which are more common in one age group (e.g., sexting in adolescents, or developmental outcome in young children).
For the effects where there were multiple age groups (e.g., body composition), differences tended to be small.

We have included some additional information on this in the *Discussion* section:

`r get_revision("r1_5")`

@@

The reviewer was pointing to the following quote from the *Lancet*: "Our understanding of the benefits, harms, and risks of our rapidly changing digital landscape is sorely lacking.".
It is not that the evidence is lacking, per se, but that our *understanding* is weak.
Our understanding is lacking partly because benefits of screen time (as seen in some educational literature) and reviewed separately from harms.
Our review aims to address this (see also Reviewer 1, \comment{r1_2}).
We think synthesising the trade-offs in one review can help us understand when screen time might be associated with good/poor outcomes, helping to better understand why.

We have added this additional context to the *Introduction*:

@@

We have revised the introduction with new papers, with a particular focus on other umbrella reviews. See the below for an example of the changes:
